### 2020-04-23 bibold.tex

%
\bibitem{ridge_aa_1}
CMS Collaboration, "Centrality dependence of dihadron correlations and azimuthal anisotropy harmonics in PbPb collisions at  $\sqrt{\it{s_{\rm{NN}}}} = $2.76 TeV",   Eur. Phys J. C72 (2012).
\bibitem{ridge_aa_2}
ALICE Collaboration, "Anisotropic Flow of Charged Particles in Pb-Pb Collisions at $\sqrt{\it{s_{\rm{NN}}}} = $5.02 TeV",  Phys. Rev. Lett. 116, 132302 (2016).
\bibitem{ridge_pp_1}
CMS Collaboration, Measurement of Long-Range Near-Side Two-Particle Angular Correlations in pp Collisions at $\sqrt{\it{s}} = $13 TeV", Phys. Rev. Lett. 116, 172302 (2016).
\bibitem{ridge_pp_2}
CMS Collaboration, "Evidence for collectivity in pp collisions at the LHC", Phys. Lett. B765 (2017) 193.
\bibitem{ridge_pp_3}
ALICE Collaboration, "Investigations of anisotropic flow using multi-particle azimuthal correlations in pp, p-Pb, Xe-Xe, and Pb-Pb collisions at the LHC", arXiv:1903.01790 [nucl-ex].
\bibitem{ridge_pp_4}
PHENIX Collaboration, "Creation of quark–gluon plasma droplets with three distinct geometries", Nature Phys. 15, 214 (2019), Nature Phys. 13, 535.
\bibitem{ridge_theory_1}
H. Mäntysaari, B. Schenke, C. Shen and P. Tribedy, "Imprints of fluctuating proton shapes on flow in proton-lead collisions at the LHC", Phys. Lett. B772 (2017) 681.
\bibitem{ridge_theory_2}
W. Zhao, Y. Zhou, H. Xu, W. Deng, and H. Song, “Hydrodynamic collectivity in proton?proton collisions at 13 TeV,” Phys. Lett. B780 (2018) 495.
\bibitem{ridge_theory_3}
K. Welsh, J. Singer, and U. W. Heinz, “Initial state fluctuations in collisions between light and heavy ions,” Phys. Rev. C94 no. 2, (2016) 024919.
\bibitem{ridge_theory_4}
M. Greif, C. Greiner, B. Schenke, S. Schlichting, and Z. Xu, “Importance of initial and final state effects for azimuthal correlations in p+Pb collisions,” Phys. Rev. D96 no. 9, (2017) 091504.
\bibitem{ALICEdet}
ALICE Collaboration, “The ALICE experiment at the CERN LHC,” JINST 3 (2008) S08002.
\bibitem{ITSpaper}
ALICE Collaboration, "Performance of the present ALICE Inner Tracking System and studies for the upgrade",  Journal of Instrumentation, 7, June (2012) C06007.
\bibitem{VZEROpaper}
ALICE Collaboration, "Performance of the ALICE VZERO system", JINST 8 (2013) P10016.
\bibitem{TPCpaper}
ALICE Collaboration, "Performance of the ALICE Time Projection Chamber", Physics Procedia 37, (2012) 434.