### 2020-06-26 paper.tex
final version to IRC first round

\documentclass[ALICE,manyauthors]{cernphprep}

\usepackage[comma,square,numbers,sort&compress]{natbib}
\usepackage{hyperref}
\usepackage{lineno}
%\usepackage{subcaption}
\usepackage{subfigure}
\usepackage[section]{placeins}
\usepackage {multicol}% Multi column in the table
\usepackage {multirow}% Multi row in the table
\usepackage {units}
\usepackage {array}
\linenumbers

\input{commands.tex}
\begin{document}

%%%%%%%%%%%%%%%  Title page %%%%%%%%%%%%%%%%%%%%%%%%
\begin{titlepage}

\PHyear{}
\PHnumber{2019-xxx}      % required, will be obtained from PH
%\PHdate{Day Month}  % required, will be obtained from PH
\PHdate{\today}
%

%%% Put your own title + short title here:
\title{Long range correlations and their event scale dependence in high multiplicity pp collisions at $\sqrt{s} = 13$~TeV}
\ShortTitle{Long-range correlations in pp collisions}   % appears on right page headers

%%% Do not change the next lines
%\Collaboration{ALICE Collaboration\thanks{See Appendix~\ref{app:collab} for the list of collaboration members}}
\ShortAuthor{ALICE Collaboration} % appears on left page headers, do not change

\begin{abstract}
%\input{abstract.tex}

Long-range azimuthal correlations are measured in high-multiplicity pp collisions at center-of-mass energy $\sqrt{s} = 13$~TeV by ALICE at LHC. 
These results are obtained in a relative pseudorapidity range of $1.6 < |\eta| < 1.8$ and transverse momentum region of $1 < p_{\mathrm T} < 5$~GeV/$c$ for the trigger and associated particles.
In addition to the conventional long-range correlation studies, the results are further investigated in events, where jets with a harder fragmentation are present.
The jets are reconstructed by using the anti-$k_\mathrm{T}$ algorithm with resolution parameter $R=0.4$ in $|\eta_\mathrm{jet}|<0.4$. 
The magnitude of the correlation dubbed ridge yield, is consistent with the results from other LHC experiments for unbiased events. 
These correlations also persist in the events containing jets. Furthermore data show that the ridge yields tend to increase in events holding $\ptjet > 10$~GeV/$c$.
The results are compared to $\pythiashoving$ calculations.  
$\pythiashoving$ model can describe low $p_{\mathrm T}$  ridge yields but can't give good description of $p_{\mathrm T}$ dependence both for the unbiased and jet events.
The results open new aspects of studying the small collision systems with jet tagged events in the future and will help to explore possible physical origins of long-range correlations.

\end{abstract}

\end{titlepage}

\setcounter{page}{2}

\input{sec_intro.tex}
\input{sec_ana.tex}
\input{sec_results.tex}
\input{sec_summary.tex}

%%%%% acknowledgements
\newenvironment{acknowledgement}{\relax}{\relax}
%\begin{acknowledgement}
%\section*{Acknowledgements}
%\input{acknowledgements.tex}    %%%%%%% done by webmaster team
%\end{acknowledgement}

\newpage
\appendix

\clearpage

\bibliographystyle{utphys}
\bibliography{paper.bib}

%%%%%%%%% appendix with author list
%\section{The ALICE Collaboration}
%\label{app:collab}
%\input{Alice_Authorlist_2017-Aug-21.tex}  %%%%%%% done by webmaster team
%\input{}

\end{document}