### 2020-08-10 paper.tex
First actions after IRC's comments

\documentclass[ALICE,manyauthors]{cernphprep}

\usepackage[comma,square,numbers,sort&compress]{natbib}
\usepackage{hyperref}
\usepackage{lineno}
%\usepackage{subcaption}
\usepackage{subfigure}
\usepackage[section]{placeins}
\usepackage {multicol}% Multi column in the table
\usepackage {multirow}% Multi row in the table
\usepackage {units}
\usepackage {array}
\linenumbers

\input{commands.tex}
\begin{document}

%%%%%%%%%%%%%%%  Title page %%%%%%%%%%%%%%%%%%%%%%%%
\begin{titlepage}

\PHyear{}
\PHnumber{2019-xxx}      % required, will be obtained from PH
%\PHdate{Day Month}  % required, will be obtained from PH
\PHdate{\today}
%

%%% Put your own title + short title here:
\title{Long-range correlations and their event-scale dependence in high-multiplicity pp collisions at $\sqrt{s} = 13$~TeV}
\ShortTitle{Long-range correlations in pp collisions}   % appears on right page headers

%%% Do not change the next lines
%\Collaboration{ALICE Collaboration\thanks{See Appendix~\ref{app:collab} for the list of collaboration members}}
\ShortAuthor{ALICE Collaboration} % appears on left page headers, do not change

\begin{abstract}
%\input{abstract.tex}

Long-range azimuthal correlations are measured in high-multiplicity pp collisions at center-of-mass energy $\sqrt{s} = 13$~TeV by ALICE at the LHC. The measurement was carried out for trigger and associated particles having transverse momenta in the range $1 < p_{\mathrm T} < 5$~GeV/$c$. Trigger and associated particles were separated by a large pseudorapidity interval $1.6 < |\Delta \eta |<1.8$.
%These results are obtained in a relative pseudorapidity range of $1.6 < |\eta| < 1.8$ and transverse momentum region of  for the trigger and associated particles.
In addition to the conventional long-range correlation studies, the results are further investigated in events with a high-pT leading track or jet.
%where jets with a harder fragmentation are present.
The jets are reconstructed by using the anti-$k_\mathrm{T}$ algorithm with resolution parameter $R=0.4$ in $|\eta_\mathrm{jet}|<0.4$. 
%The magnitude of the correlation dubbed ridge yield, is consistent with the results from other LHC experiments for unbiased events.
Measured ridge yield is consistent with the results from the CMS experiments for unbiased high-multiplicity events.
These long-range correlations are also seen in the events containing jets. Furthermore, data show that the ridge yields tend to grow with transverse momentum of the jet or leading track. %increase in events holding $\ptjet > 10$~GeV/$c$.
The results are compared to $\pythiashoving$ calculations.  
$\pythiashoving$ model can describe low $p_{\mathrm T}$ ridge yields up to 1.5 GeV/$c$ but can not give good description of $p_{\mathrm T}$ dependence both for the unbiased and jet-tagged events.
The results open new aspects of studying the small collision systems with jet tagged events in the future and will help to explore possible physical origins of long-range correlations.

\end{abstract}

\end{titlepage}

\setcounter{page}{2}

\input{sec_intro.tex}
\input{sec_ana.tex}
\input{sec_results.tex}
\input{sec_summary.tex}

%%%%% acknowledgements
\newenvironment{acknowledgement}{\relax}{\relax}
%\begin{acknowledgement}
%\section*{Acknowledgements}
%\input{acknowledgements.tex}    %%%%%%% done by webmaster team
%\end{acknowledgement}

\newpage
\appendix

\clearpage

\bibliographystyle{utphys}
\bibliography{paper.bib}

%%%%%%%%% appendix with author list
%\section{The ALICE Collaboration}
%\label{app:collab}
%\input{Alice_Authorlist_2017-Aug-21.tex}  %%%%%%% done by webmaster team
%\input{}

\end{document}