### 2020-10-24 sec_results.tex

% !TEX root = paper.tex

\section {Results}
\label{sec:results}

\subsection{Ridge yield}

\begin{figure}[h!]
	\centering
	\subfigure{ \includegraphics[width=0.47 \textwidth]{./figures/corr1.eps} }
	\subfigure{ \includegraphics[width=0.47 \textwidth]{./figures/corrmb.eps} }
	\caption{ Dihadron correlation functions as functions of $\Delta\eta$ and $\Delta\varphi$ in the high-multiplicity (0--0.1\%, left) and the minimum-bias events (0--100\%, right). The intervals of $\pttrig$ and $\ptassoc$ are equally $1 < \it{p}_{\rm{T}} < 2$ GeV/$c$. }
	\label{fig:PlotCorrMBHMT}
\end{figure}

Figure~\ref{fig:PlotCorrMBHMT} shows two-particle correlation functions as a function of the $\Delta \eta$ and $\Delta \varphi$ separation of trigger and associated particles. The per trigger yield iss measured with Eq.~\ref{eq:corrfunction} for $1 < \pttrig~\mathrm (\ptassoc) < 2$ GeV/$c$ in pp collisions at $\sqrt{\it{s}} = $\unit{13} {\rm{}TeV}. The left panel is for the high-multiplicity class. The right one is from the minimum bias events. The ridge structure is clearly observed in the high-multiplicity class while it is not seen in the minimum bias events. The away-side yield is populated mostly by correlation of jets which are back-to-back.

\begin{figure}[h!]
	\centering
	\includegraphics[width=0.9\linewidth]{./figures/Fig2_PlotDeltaPhi.eps}
	\caption{One-dimensional $\Delta\varphi$ distribution in the large $\Delta\eta$ projection for various transverse momentum intervals in each multiplicity class. The distributions in upper panels are measured with 0--100\% multiplicity class. The distribution in lower panels are measured with 0--0.1\% multiplicity class. Transverse momentum ranges of trigger particle and associated particle are $1<\pt<2$ (left), $2<\pt<3$ (middle) and $3<\pt<4$ GeV/$c$ (right), respectively. The model predictions are presented as colored lines. Blue line corresponds to $\pythiashoving$, orange line corresponds to $\pythiam$ and green line corresponds to $\epos$.}
	\label{fig:PlotDeltaPhi}
\end{figure}

Projected $\Delta\varphi$ distributions of the two-particle correlation functions in $1.6<|\Delta\eta|<1.8$ are shown in Fig.~\ref{fig:PlotDeltaPhi} for various $\it{p}_{\rm{T}}$ intervals in the minimum bias class and the high-multiplicity class. The near-side ($\Delta\varphi\sim 0$) peak in high-multiplicity class is clearly observed in all transverse momentum ranges studied while there is no hint of the signal in the minimum bias class. The near-side peak is highest in the interval of $1<\it{p}_{\rm{T}}<2$ GeV/$c$ and gradually decreases with increasing $\it{p}_{\rm{T}}$ in the high-multiplicity class. The measurements in the high-multiplicity class are compared to the results measured by the CMS collaboration~\cite{Khachatryan:2015lva}. The particle multiplicity in the present analysis is estimated with the forward V0 subsystem, whereas whereas the analysis \cite{Khachatryan:2015lva}  used the mid-rapidity particles in $|\eta|<$2.4 and $\pt>0.4$ GeV/$c$. By interpolating the $\eta$ and $\pt{}$ distributions based on PYTHIA, it is found that the multiplicity in \cite{Khachatryan:2015lva} is about 20\% higher than 0--0.1\% in this analysis, corresponding 0--0.03\%. The near-side peaks in all transverse momentum ranges are comparable to each other within the uncertainties. The larger away yields observed in \cite{Khachatryan:2015lva} can be attributed to the difference in $\eta$ acceptances. The model predictions are presented for the measurements in high-multiplicity class with the identical charged track acceptance and the $\Delta\eta$ projection range to those used in this article. The high-multiplicity class for each model is determined by requiring the minimum number of charged particles larger than 105, which corresponds to the top 0--0.1\% multiplicity class based on $\pythiam$. The $\pythiashoving$ gives good estimates of the near-side peak and slightly overestimates the away-side yield for the interval of 1$<\it{p}_{\rm{T}}<$2 GeV/$c$. However the $\pythiashoving$ underestimates the near-side peak for the $\it{p}_{\rm{T}}>$2 GeV/$c$ range. $\pythiam$ does not show any peaks in the near-side as expected. It slightly underestimates the away-side peak for $1<\it{p}_{\rm{T}}<$2 GeV/$c$ and gives good estimates for the $\it{p}_{\rm{T}}>$2 GeV/$c$ range. On the other hand, $\epos$ describes better $\pt$ dependence quantitatively for the 1$<\it{p}_{\rm{T}}<$4 GeV/$c$ range, while overestimating the near-side peak for the $\it{p}_{\rm{T}}<$2 GeV/$c$ range.

\begin{figure}[h!]
	\centering
	\subfigure{\includegraphics[width=0.99\textwidth]{./figures/Fig3_PlotRidgeYield.eps}}
	\caption{ Fully corrected near-side ridge yield as a function transverse momentum. The filled circles denote the measurement with ALICE. The CMS measurement~\cite{Khachatryan:2015lva} is represented by open blue box markers. The three lines show model predictions from $\pythiam$ (blue line), $\pythiashoving$ (orange line) and $\epos$ (green line).}
	\label{fig:PlotYSpect}
\end{figure}

Fig.~\ref{fig:PlotYSpect} shows the near-side ridge yield measured in high-multiplicity events as a function of transverse momentum. The ALICE data are compared to \cite{Khachatryan:2015lva}.
%The ridge yields as a function of the transverse momentum are shown in Fig.~\ref{fig:PlotYSpect} in the high-multiplicity class and compared to \cite{Khachatryan:2015lva}.
Taking into account the differences in acceptance of charged tracks and estimated multiplicity, the measurements are comparable with each other. The spectrum is also compared with model calculations. $\pythiam$ gives zero yields since it does not contain the ridge effect. $\pythiashoving$ describes the yield qualitatively however its yield decreases more rapidly than the measured data as $\pttrigassoc$ increases. $\epos$, unlike $\pythiashoving$, gives estimates for the $\pt$ dependence showing the ridge yield for the $\it{p}_{\rm{T}}>2$ GeV/$c$ range, while estimating larger yield for the $\it{p}_{\rm{T}}<2$ GeV/$c$.

\subsection{Event-scale dependence of the ridge yield}
\begin{figure}[h!]
	\centering
	\subfigure{ \includegraphics[width=0.47 \textwidth]{./figures/corrlh.eps} }
	\subfigure{ \includegraphics[width=0.47 \textwidth]{./figures/corrjet.eps} }
	\caption{ Two-dimensional correlations function as a function of $\Delta\eta$ and $\Delta\varphi$ in top 0-0.1\% multiplicity class with the minimum $\ptlead$ (left) or $\ptjet$ (right) selection. The interval of $\pttrig$ and $\ptassoc$ is $1<\pt<2$ GeV/$c$. The minimum requirement for $\ptlead$ or $\ptjet$ is 9 or 10 GeV/$c$, respectively. }
	\label{fig:PlotCorrHMTSel}
\end{figure}

The ridge yield is further studied with respect to two different event-scales of the hard processes. Event-scale measurement by the largest $\pt$ of charged particles in each event is denoted as $\ptlead$ while the other is done by reconstructing jets and measuring their transverse momenta, represented as $\ptjet$. As shown in Fig.~\ref{fig:PlotCorrHMTSel}, the ridge structure for $1< \pttrig\,(\ptassoc) <2$ GeV/$c$ still persists in high-multiplicity pp collisions with $\ptlead>9$~GeV/$c$ (left) or $\ptjet>10$~GeV/$c$ (right) .  %The interval is the same to the measured ridge yield is maximum without the event-scale selection.
It is worth noting that the correlation function with the minimum $\ptjet$ selection has double peaks along the $\Delta\eta$ axis at $\Delta\varphi = \pi$ due to the jet tagging with the limited jet acceptance in the mid-rapidity as described in this article.

\begin{figure}[h!]
	\centering
	\includegraphics[width=0.99\linewidth]{./figures/Fig5_PlotDeltaPhiESE.eps}
	\caption{ One-dimensional $\Delta\varphi$ distribution in the large $\Delta\eta$ with the minimum $\ptlead$(lower) or $\ptjet$(upper) selection for $1< \pttrig~\mathrm (\ptassoc) <2$ GeV/$c$ in the 0--0.1\% multiplicity class. The filled circles show the ALICE measurement. Model predictions are presented with a blue line for $\pythiam$, an orange line for $\pythiashoving$ and a green line for $\epos$.}
	\label{fig:PlotDeltaPhiESE}
\end{figure}

Figure~\ref{fig:PlotDeltaPhiESE} shows projected $\Delta\varphi$ distributions of the correlation functions in $1.6<|\Delta\eta|<1.8$ with the minimum $\ptlead$ (lower) and $\ptjet$ (upper) requirement. Even with the event-scale selection, the ridge is still visible at the near-side. The near-side yield increases as the threshold of $\ptjet$ increases compared to one measured in unbiased events, while the ridge yields for an increasing threshold of $\ptlead$ increase less than those of $\ptjet$. \iffalse The event-scale measurement with $\ptlead$ favors harder jets as a leading particle carries a large fraction of the momentum. In contrast, soft jets are favored for reconstructed jets. \fi The results are compared with $\pythiashoving$, $\pythiam$, and $\epos$ calculations. The near-side peaks are qualitatively reproduced by $\pythiashoving$ and $\epos$, while underestimated by $\pythiashoving$ and overestimated by $\epos$. On the other hand, $\pythiam$ does not show the near-side peak with the event-scale selections, while giving similar estimates for away-side yield as $\pythiashoving$.
%Both models describe the away-side yield with the $\ptjet$ selection and overestimate away-side yield with the $\ptlead$ selection.

\begin{figure}[h!]
	\centering
	\includegraphics[width=0.99\linewidth]{./figures/Fig6_RidgeYieldESE.eps}
	\caption{The ridge yield spectra with respect to the leading particle and jet selections. The filled circles show measurement with ALICE. Model predictions are compared to measurement as blue lines for $\pythiam$, orange lines for $\pythiashoving$ and green lines for $\epos$. Ratios of measured data to models are prepared for $\pythiashoving$ and $\epos$.}
	\label{fig:RidgeYield_ESE}
\end{figure}

Finally, the ridge yields as functions of the minimum $\ptlead$ ($\it{p}^{\rm{LP}}_{\rm{T,min}}$) and $\ptjet$ ($\it{p}^{\rm{Jet}}_{\rm{T,min}}$) selections are shown in Fig.~\ref{fig:RidgeYield_ESE}. The ridge yields are increased for $\it{p}^{\rm{LP}}_{\rm{T,min}} >$ 7 and $\it{p}^{\rm{Jet}}_{\rm{T,min}} \ge$ 10~GeV/$c$ compared to one from the unbiased events. A moderate increase of the ridge yields toward higher $\ptlead$ or $\ptjet$ is observed and there is no difference between two event selections within the uncertainties.
The results are compared with $\pythiashoving$, $\pythiam$ and $\epos$. $\pythiashoving$ describes comparable trend with data, underestimating the ridge yield. On the other hand, $\epos$ also describes comparable trend with data, overestimating the ridge yield. The origin of the enhanced ridge yields for higher momentum jet-tagged events is not clear to date but it might be related with the expected shorter impact parameters for dijet or multi-jet production events studied in~\cite{Frankfurt:2010ea}.