### 2020-12-06 sec_intro.tex

% !TEX root = paper.tex

\section{Introduction}
\label{sec:intro}

In high-energy nucleus-nucleus collisions at RHIC~\cite{Adams:2005dq,Adcox:2004mh,Arsene:2004fa,Back:2004je} and LHC~\cite{Abelev:2012di, Abelev:2014pua, ATLAS:2011ah}, significant correlations are observed between particles emitted over a wide pseudorapidity range. The origin of these observations are collective effects, which are related to the formation of a strongly interacting matter called the quark-gluon plasma (QGP), which exhibits hydrodynamic behavior (see the reviews~\cite{Romatschke:2007mq,Jeon:2015dfa,Romatschke:2017ejr}). 
Recent theoretical~\cite{Niemi:2015qia,Bernhard:2016tnd,Bernhard2019} and experimental~\cite{ALICE:2016kpq,Acharya:2017gsw,Acharya:2017zfg,Acharya:2020taj} advancements have contributed significantly to understanding the transport properties of the QGP.
Similar long-range correlations, typically interpreted as signatures of collectivity in heavy-ion collisions, are also observed in high-multiplicity proton-proton (pp)~\cite{Aad:2015gqa,Khachatryan:2015lva,Khachatryan:2016txc,Acharya:2019vdf}, proton-nucleus (pA)~\cite{Abelev:2012ola,Aad:2014lta,Aaboud:2016yar,Khachatryan:2016ibd}, and lighter nucleus-nucleus systems~\cite{PHENIX:2018lia,Aidala:2017ajz}.
Azimuthal correlations of particles separated by a large gap in pseudorapidity may suggest hydrodynamic behavior even in these small systems, although the volume and lifetime of the medium produced by such a colliding system are expected to be small, and there are other mechanisms which can produce similar flow-like signals~\cite{Busza:2018rrf,Nagle:2018nvi}.
%The observed long-range azimuthal correlations suggest flow with hydrodynamic characteristics even in these small systems, though the volume and lifetime of the medium produced by such a system are expected to be small~\cite{Busza:2018rrf,Nagle:2018nvi}.

Measurements of two-particle angular correlations give information on many physical effects, including collectivity, hadronization, fragmentation, and femtoscopic effects~\cite{Lisa:2005dd}, and are typically quantified as a function of $\Delta\eta$, the relative pseudorapidity, and $\Delta\varphi$, the separation in azimuthal angle, of pairs of particles. The long-range structure of two-particle correlation functions is well suited to analyse collective effects, since it is not created by resonance decays and fragmentation of high-momentum partons. A typical source of long-range correlations in Monte Carlo pp generators is momentum conservation. %and away-side ($\Delta\varphi$ $\approx$ $\pi$) jet correlations.
The enhancement in the associated yield of two-particle correlations at small $\Delta\varphi$ that extends over a large $\Delta\eta$, is dubbed ``ridge'' due to its characteristic shape in the $\Delta\eta$--$\Delta\varphi$ plane.
The shape of these $\Delta\varphi$ correlations can be studied via a Fourier decomposition~\cite{Poskanzer:1998yz,Voloshin:2008dg}. The second and third order terms are the dominant harmonic coefficients. In heavy-ion collisions, harmonic coefficients can be related to the collision geometry and density fluctuations of the colliding nuclei~\cite{Alver:2010gr,Alver:2010dn,ALICE:2011ab} and to transport properties of the QGP in relativistic viscous hydrodynamic models~\cite{Gale:2012rq,Niemi:2015qia,Shen:2014vra,Bernhard:2016tnd,Bernhard2019}.

% TODO more pA measurements
The ridge structures in high-multiplicity pp and p--Pb events have been attributed to initial-state effects or final-state effects. The initial-state effects usually referred to gluon saturation~\cite{Dusling:2012cg,Bzdak:2013zma} and color reconnections~\cite{Ortiz:2013yxa,Sarma:2019teo}, which are formed along the longitudinal direction. The final-state effects may contribute via parton-induced interactions~\cite{Arbuzov:2011yr} or collective effects due to hydrodynamic behavior of the produced particles arising in a high-density system possibly formed in these collisions~\cite{Weller:2017tsr,Zhao:2017rgg}. 
Hybrid models implementing both effects are generally used in hydrodynamic simulations~\cite{Greif:2017bnr,Mantysaari:2017cni}. 
The proton shape and its fluctuations are also important to model small systems~\cite{Mantysaari:2017cni}.
%Nevertheless, the hydrodynamics itself might not be the only mechanism of the observed collectivity~\cite{Zhao:2020pty,Schenke:2019pmk}. 
To understand the influence of initial or final state effects, and possibly disentangle the two, a quantitative description of the measurements in small systems~\cite{Schenke:2019pmk,Schenke:2020mbo} needs to account for details of the initial state.
Systematic studies of these correlation effects from small to large systems are being performed, both experimentally~\cite{Acharya:2019vdf} and theoretically~\cite{Schenke:2020mbo}.
However, the quantitative description of the full set of experimental data has not been achieved yet.
A summary of various explanations for the observed correlations in small systems is given in ~\cite{Strickland:2018exs,Loizides:2016tew,Nagle:2018nvi}.

Besides the hybrid models mentioned above, alternative approaches have been developed to describe collectivity in small systems. A microscopic model for collectivity has been implemented in the PYTHIA8 event generator, which is based on interacting strings and is called the “string shoving model”~\cite{Bierlich:2017vhg}. In this model, strings repel each other in the transverse direction, which results in microscopic transverse pressure and, consequently, in long-range correlations. PYTHIA with string shoving can qualitatively reproduce the CMS near-side ridge yield~\cite{Khachatryan:2016txc}. This challenges the hydrodynamic picture and predicts modifications to jet fragmentation properties~\cite{Bierlich:2019ixq}.
EPOS LHC describes collectivity in small systems with a parameterized hydrodynamic evolution of the high-energy density region, so called ``core'', formed by many color string fields~\cite{Pierog:2013ria}.

If collectivity in small systems is due to final-state interactions, it is expected 
that these final-state interactions will affect also produced jets. Proving the presence of jet quenching would be another crucial evidence of the existence of a high-density strongly-interacting system, possibly a  QGP, in high-multiplicity pp collisions. However, so far there has been no sign of jet quenching observed in high-multiplicity pp and p--Pb collisions~\cite{Khachatryan:2016odn,Adam:2016jfp,Adam:2016dau,Acharya:2017okq}. Jet fragmentation can be studied in two-particle correlation functions by looking at short-range correlations around $(\Delta\eta$,$\Delta\varphi)=(0,0)$~\cite{Adam:2016tsv}.  
%The choice of a reference system can thus be ambiguous especially for pp since the hard probes themselves are enhanced by requiring high-multiplicity event~\cite{Adam:2016jfp,Acharya:2018egz}, also resulting in enhancement of probability to select events with high-$\pt$ jets.
%Small system studies: A theory overview 1807.07191 Strickland:2018exs
%find one more one this regard.
%The ATLAS experiment recently showed that the ridge remains in events tagged with a Z-boson~\cite{Aaboud:2019mcw}, possibly with an accompanying jet.
%The impact parameter dependence on dijet or multi-jet production in pp collisions was studied in~\cite{Frankfurt:2003td,Frankfurt:2010ea}.

% check this paper https://arxiv.org/abs/1910.13978 ATLAS pt cut...
% Bierlich:2017vhg (Shoving) Bierlich:2019ixq  
% add vn extractio??
% flow vn extraction : awayside ??
% 
% AA geometry dominant for v2 and what about in pp, b vs v2 or ridge..
% Frankfurt:2003td
% Aaboud:2019mcw Z-tagged ridge ATLAS

To further investigate the interplay of jet production and collective effects in small systems, we simultaneously study long- and short-range correlations in very-high-multiplicity pp collisions at $\sqrt{s} =13$ TeV, collected with the high-multiplicity event trigger during the LHC Run 2 data taking period. In this article, we present the near-side per-trigger yield at large pseudorapidity separation as a function of transverse momentum. The results are compared to the previous measurements by the CMS collaboration~\cite{Khachatryan:2015lva}. %In addition, we report the ridge yield in events, where a hard scattering took place, to explore possible physical origins of the long-range correlations.
In addition, we report the ridge yield and near-side jet-like correlations with event-scale selection. The event-scale selection is done by requiring a minimum transverse momentum of the leading track or the reconstructed jet at mid-rapidity, which is expected to bias the impact parameter of pp collisions to be smaller on average ~\cite{Sjostrand:1986ep,Frankfurt:2010ea}.
The experimental setup and analysis method are described in Sec.~\ref{sec:experiment} and \ref{sec:ana}, respectively. The sources of systematic uncertainties are discussed in Sec.~\ref{sec:uncertainties}. The results of the measurements and comparisons to model calculations are presented in Sec.~\ref{sec:results}. Finally, Sec.~\ref{sec:summary} summarizes the obtained results.

 %h has indicated the formation of a strongly interacting quark-gluon plasma (QGP) matteras indicated the formation of a strongly interacting quark-gluon plasma (QGP) matter