### 2020-12-06 sec_results.tex

% !TEX root = paper.tex

\section {Results}
\label{sec:results}

\subsection{Ridge yield}

\begin{figure}[h!]
	\centering
	\subfigure{ \includegraphics[width=0.47 \textwidth]{./figures/corr1.pdf} }
	\subfigure{ \includegraphics[width=0.47 \textwidth]{./figures/corrmb.pdf} }
	\caption{ Dihadron correlation functions as functions of $\Delta\eta$ and $\Delta\varphi$ in high-multiplicity (0--0.1\%, left) and minimum-bias events (0--100\%, right). The intervals of $\pttrig$ and $\ptassoc$ are equally $1 < \it{p}_{\rm{T}} < $2 GeV/$c$. }
	\label{fig:PlotCorrMBHMT}
\end{figure}

Figure~\ref{fig:PlotCorrMBHMT} shows the per-trigger yield obtained from Eq. (1) for $1 < \pttrig~\mathrm (\ptassoc) < $2 GeV/$c$ in pp collisions at $\sqrt{\it{s}} = $\unit{13} {\rm{}TeV} for high-multiplicity (left) and minimum bias events (right panel). The ridge structure is clearly observed in the high-multiplicity class while it is less visible and conclusive in the minimum bias events. The away-side yield is populated mostly by back-to-back jet correlations.

\begin{figure}[h!]
	\centering
	\includegraphics[width=0.9\linewidth]{./figures/Fig2_PlotDeltaPhi.pdf}
	\caption{One-dimensional $\Delta\varphi$ distribution in the large $\Delta\eta$ projection for three transverse momentum intervals in high-multiplicity (lower panels) and minimum bias (upper panels) events after ZYAM subtraction. Transverse momentum ranges of trigger particles and associated particles are $1<\pt<$2 (left), 2$<\pt<$3 (middle) and 3$<\pt<$4 GeV/$c$ (right), respectively. The presented model predictions were calculated using $\pythiashoving$, $\pythiam$, and $\epos$.}
	\label{fig:PlotDeltaPhi}
\end{figure}

Figure~\ref{fig:PlotDeltaPhi} shows $\Delta\varphi$ projections of the two-particle correlation functions obtained in the range 1.6$<|\Delta \eta |<$1.8 for several track $\pt$ intervals after ZYAM subtraction, see Eq. (2). Results are shown for various $\it{p}_{\rm{T}}$ intervals in the minimum bias class (upper) and the high-multiplicity class (lower) down to 1 GeV/$c$ where non-flow contamination is negligible. The near-side ($\Delta\varphi\sim 0$) ridge in the high-multiplicity class is clearly observed in all transverse momentum ranges studied while there is no definitive signal in the minimum bias class. Within the range of analyzed particle $\pt$, the yield in the near-side ridge decreases with increasing $\pt$ in the high-multiplicity class. The measurements in the high-multiplicity class are compared to the results published by the CMS collaboration~\cite{Khachatryan:2015lva}. In case of the CMS measurement, the charged particle multiplicity was obtained by counting $\pt>0.4$ GeV/$c$ tracks in $|\eta|<$2.4. In our analysis, event multiplicity is determined from the forward V0 detectors. The difference in multiplicity selection between ALICE (forward) and CMS (mid-rapidity) is studied using PYTHIA 8 simulations and it is found that the CMS multiplicity is ${\approx}20\,\%$ larger than the ALICE one when compared in the acceptance region of the measurements reported here, $|\eta|<$0.9. Near-side ridges in all transverse momentum ranges are comparable to each other within the uncertainties. The larger away-side yields observed in \cite{Khachatryan:2015lva} can be attributed to the difference in $\eta$ acceptance for the multiplicity selection, which overlaps somewhat with the $\eta$ region in Fig.~\ref{fig:PlotDeltaPhi} for the CMS results. The ALICE data are compared with model predictions where we apply a comparable high-multiplicity selection and $\Delta\eta$ projection range. Selection of high-multiplicity events in models is done by requiring a minimal number of charged particles that pass the V0M acceptance. This threshold is set to 105, which is tuned to reach the top 0--0.1\% centrality selection using $\pythiam$. The threshold for $\epos$ or $\pythiashoving$ is 110 or 108, respectively. The statistical uncertainties due to the limited number of events for model calculations are shown as band in each figure. The $\pythiashoving$ gives good estimates of the near-side ridge yield and slightly overestimates the away-side yield for the interval of 1$<\it{p}_{\rm{T}}<$2 GeV/$c$. However, the $\pythiashoving$ underestimates the near-side ridge yield for $\it{p}_{\rm{T}}>$2 GeV/$c$. $\pythiam$ does not show any near-side ridge as expected. It slightly underestimates the away-side peak for $1<\it{p}_{\rm{T}}<$2 GeV/$c$ and gives good estimates for $\it{p}_{\rm{T}}>$2 GeV/$c$. On the other hand, $\epos$ describes quantitatively better the $\pt$ dependence in the 2$<\it{p}_{\rm{T}}<$4 GeV/$c$ range, while overestimating the near-side ridge yield for the $\it{p}_{\rm{T}}<$2 GeV/$c$ range.

\begin{figure}[h!]
	\centering
	\subfigure{\includegraphics[width=0.99\textwidth]{./figures/Fig3_PlotRidgeYield.pdf}}
	\caption{ Fully corrected near-side ridge yield as a function transverse momentum. The filled circles denote the measurement with ALICE. The CMS measurement~\cite{Khachatryan:2015lva} is represented by open blue box markers and extends down to lower $\pt$ due to the larger $\Delta\eta$ acceptance. The three lines show model predictions from $\pythiam$ (blue line), $\pythiashoving$ (orange line) and $\epos$ (green line).}
	\label{fig:PlotYSpect}
\end{figure}

Figure~\ref{fig:PlotYSpect} shows the near-side ridge yield measured in high-multiplicity events as a function of transverse momentum. The ALICE data are compared with results from CMS~\cite{Khachatryan:2015lva}.
%The ridge yields as a function of the transverse momentum are shown in Fig.~\ref{fig:PlotYSpect} in the high-multiplicity class and compared to \cite{Khachatryan:2015lva}.
Taking into account the differences in acceptance of both measurements and in the chosen multiplicity estimator, a perfect agreement between the two sets of results is not expected. The data is also compared with model calculations. $\pythiam$ gives zero ridge yields since it does not account for the ridge effect. $\pythiashoving$ describes the yield qualitatively, however the predicted yield decreases more rapidly than the measured one as $\pttrigassoc$ increases. $\epos$, unlike $\pythiashoving$, describes well the $\pt$ dependence of the ridge yield for the $\it{p}_{\rm{T}}>$2 GeV/$c$ range, while predicting larger yield for $\it{p}_{\rm{T}}<$2 GeV/$c$.

\subsection{Event-scale dependence of the ridge yield}
\begin{figure}[h!]
	\centering
	\subfigure{ \includegraphics[width=0.47 \textwidth]{./figures/corrlh.pdf} }
	\subfigure{ \includegraphics[width=0.47 \textwidth]{./figures/corrjet.pdf} }
	\caption{ Two-dimensional correlation functions as a function of $\Delta\eta$ and $\Delta\varphi$ in high-multiplicity events where a selection on event-scale was required in addition. The interval of $\pttrig$ and $\ptassoc$ is 1$<\pt<$2 GeV/$c$. Left: HM events with a $\ptlead>$ 9 GeV/$c$ leading track. Right: HM events with a $\ptjet>$ 10 GeV/$c$.}
	\label{fig:PlotCorrHMTSel}
\end{figure}

The ridge yield is further studied with respect to two different event-scales. In the first measurement, the event-scale is set by requiring a minimum $\pt$ cutoff on the leading track in event (denoted as $\ptlead$), while in the second measurement, we impose a minimum $\pt$ cutoff on the leading jet (denoted as $\ptjet$). As shown in Fig.~\ref{fig:PlotCorrHMTSel}, the ridge structure for $1< \pttrig\,(\ptassoc) <2$ GeV/$c$ still persists in high-multiplicity pp collisions with $\ptlead>9$~GeV/$c$ (left) and $\ptjet>10$~GeV/$c$ (right).  %The interval is the same to the measured ridge yield is maximum without the event-scale selection.
It is worth noting that the correlation function obtained with the minimum $\ptjet$ selection has a double peak structure which is oriented along the $\Delta\eta$ axis at $\Delta\varphi=\pi$. This structure emerges due to the restricted acceptance of jet tagging, $|\eta_{\rm{jet}}|<$0.4.

\begin{figure}[h!]
	\centering
	\includegraphics[width=0.99\linewidth]{./figures/Fig5_PlotDeltaPhiESE.pdf}
	\caption{ $\Delta\varphi$ projections of the correlation functions constrained to 1.6$<|\Delta\eta|<$1.8 in HM events with an additional event-scale bias. Top: with an imposed cut on leading jet $\pt$,  bottom: with an imposed cut on leading hadron $\pt$. ALICE data are compared with prediction of models.}
	\label{fig:PlotDeltaPhiESE}
\end{figure}

Figure~\ref{fig:PlotDeltaPhiESE} shows projected $\Delta\varphi$ distributions of the correlation functions in $1.6<|\Delta\eta|<1.8$ with the minimum $\ptlead$ (lower) and $\ptjet$ (upper) requirement. Even with the event-scale selection, the ridge is still visible at the near-side. The near-side ridge peak increases as the thresholds of $\ptlead$ and $\ptjet$ increase compared to the one measured in unbiased events. The results are compared with $\pythiashoving$, $\pythiam$, and $\epos$ calculations. Near-side ridge peaks are qualitatively reproduced by $\pythiashoving$ and $\epos$. On the other hand, $\pythiam$ does not show the  near-side ridge peak for neither of the two event-scale selections, but it gives compatible results for the away-side yield as $\pythiashoving$.
%Both models describe the away-side yield with the $\ptjet$ selection and overestimate away-side yield with the $\ptlead$ selection.

\begin{figure}[h!]
	\centering
	\includegraphics[width=0.89\linewidth]{./figures/Fig6_RidgeYieldESE.pdf}
	\caption{Near-side ridge yield as a function of the $\ptlead$ (left) and $\ptjet$ (right) bias. Data points (filled circles) show the ALICE measurement. They are compared to predictions of models which are represented by color bands. The bottom panel shows a ratio of the models to the data. The total uncertainty on the data is represented by the gray band centered around unity.}
	\label{fig:RidgeYield_ESE}
\end{figure}

The ridge yield as function of the minimum $\ptlead$ ($\it{p}^{\rm{LP}}_{\rm{T,min}}$) and $\ptjet$ ($\it{p}^{\rm{Jet}}_{\rm{T,min}}$) selections are shown in Fig.~\ref{fig:RidgeYield_ESE}. High-multiplicity events with imposed event-scale bias exhibit increased ridge yields when compared to unbiased HM events. A moderate increase of the ridge yields as a function of $\ptlead$ or $\ptjet$ is observed and there is no difference between the two event selections within the uncertainties.
Comparison to model calculations suggests that $\pythiashoving$ gives comparable trend with data, but underestimates the ridge yield. On the other hand, $\epos$, which also provides a trend comparable with the data, overestimates the ridge yield. The origin of the enhanced ridge yields for higher momentum jet-tagged events is not clear to date but it might be related to the expected shorter impact parameters for dijet or multi-jet production events studied in~\cite{Frankfurt:2010ea}.

\begin{figure}[h!]
	\centering
	\includegraphics[width=0.89\linewidth]{./figures/Fig9_JetYieldESE.pdf}
	\caption{ Near-side jet-like peak yield with respect to the minimum leading particle (left) and jet selections (right). The filled circles show measurement with ALICE. The measurements are compared with model descriptions from $\pythiam$, $\pythiashoving$, and $\epos$ for both selections. The total uncertainty on the data is represented by the gray
band centered around unity.}
	\label{fig:JetYield_ESE}
\end{figure}

Finally, the near-side jet-like peak yield is measured as a function of minimum $\ptlead$ and $\ptjet$ in Fig.~\ref{fig:JetYield_ESE} to further test the models that aim to describe the near-side ridge.
%quantify the trend of the jet yield.
$\epos$ provides comparable estimates of the near-side jet-like peak yield, while $\pythiam$ and $\pythiashoving$ overestimate near-side yields for both selections.

In all models where the ridge is due to final-state interactions, e.g., $\epos$ and PYTHIA 8 String Shoving, one also expects the near-side jet-like peak yield to be affected. This can be observed when comparing the measured near-side jet yields with PYTHIA 8 calculations with and without String Shoving. The new ALICE results therefore provide constraints beyond traditional ridge measurements that challenge existing models.