### 2020-12-17 paper.tex
update after CR1

\documentclass[ALICE,manyauthors]{cernphprep}

\usepackage[comma,square,numbers,sort&compress]{natbib}
\usepackage{hyperref}
\usepackage{lineno}
%\usepackage{subcaption}
\usepackage{subfigure}
\usepackage[section]{placeins}
\usepackage {multicol}% Multi column in the table
\usepackage {multirow}% Multi row in the table
\usepackage {units}
\usepackage {array}
\linenumbers

\input{commands.tex}
\begin{document}

%%%%%%%%%%%%%%%  Title page %%%%%%%%%%%%%%%%%%%%%%%%
\begin{titlepage}

\PHyear{}
\PHnumber{2020-xxx}      % required, will be obtained from PH
%\PHdate{Day Month}  % required, will be obtained from PH
\PHdate{\today}
%

%%% Put your own title + short title here:
\title{Long- and short-range correlations and their event-scale dependence in high-multiplicity pp collisions at $\boldsymbol{\sqrt{{\textit s}}}=13$~TeV}
\ShortTitle{Long-range correlations in pp collisions at $\sqrt{s} = 13$~TeV}   % appears on right page headers

%%% Do not change the next lines
%\Collaboration{ALICE Collaboration\thanks{See Appendix~\ref{app:collab} for the list of collaboration members}}
\ShortAuthor{ALICE Collaboration} % appears on left page headers, do not change

\begin{abstract}
%\input{abstract.tex}
Two-particle angular correlations are measured in high-multiplicity proton--proton collisions at $\sqrt{s} =13$~TeV by the ALICE Collaboration. The yields of particle pairs at short-($\Delta\eta$ $\sim$ 0) and long-range ($1.6 < |\Delta\eta| < 1.8$) in pseudorapidity are extracted on the near-side ($\Delta\varphi$ $\sim$ 0).
They are reported as a function of transverse momentum ($p_{\mathrm T}$) in the range 1~$<p_{\mathrm T}<$~4 GeV/$c$.
Furthermore, the event-scale dependence is studied for the first time by requiring the presence of high-$p_{\rm T}$ leading particles and jets for varying $p_{\rm T}$ thresholds. 
The results demonstrate that the long-range ``ridge'' yield, possibly related to the collective behavior of the system, is present in events with high-$p_{\mathrm T}$ processes. The magnitudes of the short- and long-range yields are found to grow with the event scale. 
The results are compared to EPOS~LHC and PYTHIA~8 calculations, with and without string-shoving interactions. It is found that while both models describe the qualitative trends in the data, calculations from EPOS~LHC show a better quantitative agreement, in particular for the $p_{\rm T}$ and event-scale dependencies.

%Long-range azimuthal correlations are measured in high-multiplicity pp collisions at center-of-mass energy $\sqrt{s} = 13$~TeV by the ALICE collaboration at the LHC. The measurement was carried out for trigger and associated particles having transverse momenta in the range $1 < p_{\mathrm T} < 5$~GeV/$c$. Trigger and associated particles were separated by a large pseudorapidity interval $1.6 < |\Delta \eta |<1.8$. Measured ridge yield is consistent with the results from the CMS collaboration.
%These resfults are obtained in a relative pseudorapidity range of $1.6 < |\eta| < 1.8$ and transverse momentum region of  for the trigger and associated particles.
%In addition to the conventional long-range correlation studies, the results are further investigated in events with a high-$p_{\rm{T}}$ leading track or jet.
%where jets with a harder fragmentation are present.
%The jets are reconstructed by using the anti-$k_\mathrm{T}$ algorithm with resolution parameter $R=0.4$ in $|\eta_\mathrm{jet}|<0.4$. 
%The magnitude of the correlation dubbed ridge yield, is consistent with the results from other LHC experiments for unbiased events.
%These long-range correlations are also seen in the events containing high-$\pt$ jets for the first time. Furthermore, data show that the ridge yield tends to grow with transverse momentum of the jet or leading track. %increase in events holding $\ptjet > 10$~GeV/$c$.
%Event-scale selection is further studied by measuring the near-side jet-like peak.
%The results are compared to $\pythiashoving$ and $\epos$ calculations.  
%$\pythiashoving$ model can describe low $p_{\mathrm T}$ ridge yields up to 1.5 GeV/$c$ but can not give a good description of $p_{\mathrm T}$ dependence both for the unbiased and jet-tagged events. $\epos$ model can describe $\pt$ dependence of ridge yield while overestimating the yield at the low $\pt<$ 2 GeV/$c$. On the other hand, near-side jet-like correlations are overestimated by $\pythiashoving$ and quantitatively estimated by $\epos$, which is opposite description to the ridge yield.

\end{abstract}

\end{titlepage}

\setcounter{page}{2}

\input{sec_intro.tex}
\input{sec_ana.tex}
\input{sec_results.tex}
\input{sec_summary.tex}

%%%%% acknowledgements
\newenvironment{acknowledgement}{\relax}{\relax}
\begin{acknowledgement}
\section*{Acknowledgements}
\noindent The ALICE Collaboration would like to thank Christian Bierlich for providing the PYTHIA8 String Shoving configurations.
%{\it (TBI: The default acknowledgments will be done by EB centrally.)}
%\input{acknowledgements.tex}    %%%%%%% done by webmaster team
\end{acknowledgement}

\newpage
\appendix

\clearpage

\bibliographystyle{utphys}
\bibliography{paper.bib}

%%%%%%%%% appendix with author list
%\section{The ALICE Collaboration}
%\label{app:collab}
%\input{Alice_Authorlist_2017-Aug-21.tex}  %%%%%%% done by webmaster team
%\input{}

\end{document}